---
title: "Option B — an estimate for a defined area"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  results: "outputs/option_B/results.rds"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
r <- readRDS(params$results)
P <- r$project; S <- r$settings; res <- r$res
D <- S$REPORT_DEPTH_CM
fig <- function(f) knitr::include_graphics(file.path(r$figures, f))
tab <- function(x, ...) knitr::kable(x, row.names = FALSE, ...)
f1 <- function(x) ifelse(is.na(x), "", formatC(x, format = "f", digits = 1))
f0 <- function(x) ifelse(is.na(x), "", formatC(x, format = "f", digits = 0, big.mark = ","))
design_text <- c(
  exploratory = "**Exploratory** — the cores were not placed with known selection probabilities (for example, chosen by the team or by an earlier study). The mean and total are a reasonable first description, but there is no defensible confidence interval, so none is given.",
  srs = "**Simple random sampling** (or systematic sampling with a random start) across the whole area.",
  stratified = "**Stratified random sampling** — random placement within each stratum; strata are weighted by their area.")[[res$design]]
n_cores <- sum(r$chk$cores$status == "Complete")
```

**`r r$project$title`**

```{r banner, results = "asis"}
if (isTRUE(S$BOUNDARY_IS_HYPOTHETICAL))
  cat("> ⚠️ **Teaching example.** The reporting boundary in this report is **hypothetical** — drawn to",
      "show how the calculation works, not a mapped meadow. The area, the total and anything that",
      "depends on them describe no real place.\n")
if (grepl("SYNTHETIC", toupper(P$title)))
  cat("\n> ⚠️ **Synthetic data.** Every core in this report was generated by a computer to demonstrate",
      "the calculation. None of it is a field observation.\n")
```

## 1. The question

`r P$question_B`

## 2. Area, habitat and sampling

| | |
|---|---|
| Area and habitat | `r P$area_habitat` |
| Reporting boundary | `r basename(S$BOUNDARY_FILE)` — `r f1(r$area_m2 / 1e4)` ha`r if (isTRUE(S$BOUNDARY_IS_HYPOTHETICAL)) " (**hypothetical**)" else ""` |
| Survey dates | `r P$survey_dates` |
| Sampling design | `r P$design` |
| Sampling unit | one `r S$SAMPLING_UNIT` (`r S$PLOT_AREA_M2` m²); cores sharing a Plot ID are averaged first |
| Independent sampling units | `r nrow(r$units)` (from `r n_cores` complete cores) |
| Data source | `r P$data_source` |

`r design_text`

## 3. Carbon pool and depth

**Sediment organic carbon** only, from the surface to **`r D` cm** (in situ). This is a stock —
the carbon in the sediment now — not a rate of accumulation and not removal of carbon from the
atmosphere.

Every core is shorter than some of the depths reported, so part of each stock is **estimated**
below the base of the core. How much:

```{r pct}
ad <- r$all_depths
x <- data.frame(Depth = ad$depth, "Mean (Mg C/ha)" = f1(ad$mean_Mg_ha),
                "Measured" = paste0(f0(100 - ad$pct_estimated), "%"),
                "Estimated below the cores" = paste0(f0(ad$pct_estimated), "%"), check.names = FALSE)
tab(x, caption = "Share of the area mean that was measured and estimated, by reporting depth")
```

## 4. How it was calculated

1. **The same checked core stocks as Option A.** Every slice passed the workbook's checks; cores
   that did not are left out (section 7).
2. **A carbon curve for each core, 0–100 cm, in 1 cm steps.** Inside the measured core each cm
   takes its slice's carbon, so the curve adds up to exactly the measured stock. Below the core,
   carbon density is estimated — first with a decay curve fitted to that core, then one fitted to
   all cores in its stratum, and if neither works, the mean of the deepest 5 cm held constant.
3. **One value per sampling unit** at `r D` cm.
4. **An area estimate that matches the design** (section 2).

```{r methods}
m <- r$methods_used
m$core_base_cm <- f1(m$core_base_cm)
m$floor_mg_cm3 <- f1(m$floor_mg_cm3)
names(m) <- c("Core", "Measured to (cm)", "Estimated below the core with", "Why not the core's own curve",
              "Why not the pooled curve", "Curve floor (mg C/cm³)")
tab(m)
```

```{r fig-curves, out.width = "65%"}
fig("carbon_curves.png")
```

```{r incs}
ci <- r$cinc
ci <- data.frame(Core = ci$core_id, Increment = ci$increment, Measured = f1(ci$measured_Mg_ha),
                 Estimated = f1(ci$estimated_Mg_ha), Total = f1(ci$stock_Mg_ha),
                 "% estimated" = f0(ci$pct_estimated), check.names = FALSE)
tab(ci, caption = "Each core by standard increment (Mg C/ha)")
```

## 5. The estimate

```{r units}
u <- r$units
u <- data.frame(Unit = u$unit_id, Stratum = u$stratum, Cores = u$n_cores,
                Measured = f1(u$measured_Mg_ha), Estimated = f1(u$estimated_Mg_ha), Total = f1(u$stock_Mg_ha))
tab(u, caption = sprintf("Sampling units, 0–%g cm (Mg C/ha)", D))
```

**`r describe <- sprintf("Mean %s Mg C/ha over %s ha: about %s Mg C in the top %g cm of sediment.", f1(res$mean_Mg_ha), f1(res$area_ha), f0(res$total_Mg_C), D); describe`**
`r if (!is.null(res$ci_Mg_ha)) sprintf("%g%% interval: %s–%s Mg C/ha, or %s–%s Mg C in total (±%s%% of the mean; the planning target was ±%s%%, which was %s).", 100 * res$conf, f1(res$ci_Mg_ha[1]), f1(res$ci_Mg_ha[2]), f0(res$ci_total_Mg_C[1]), f0(res$ci_total_Mg_C[2]), f0(100 * res$relative_margin), f0(100 * res$target), if (res$relative_margin <= res$target) "met" else "not met") else if (res$design == "exploratory") sprintf("Range among the %d sampling units: %s–%s Mg C/ha. This range is descriptive; it is not a confidence interval.", res$n_units, f1(res$range[1]), f1(res$range[2])) else ""`

```{r strata}
if (!is.null(res$by_stratum)) {
  b <- res$by_stratum
  tab(data.frame(Stratum = b$stratum, "Area (ha)" = f1(b$area_ha), "Sampling units" = b$n_units,
                 "Mean (Mg C/ha)" = ifelse(b$n_units > 0, f1(b$mean_Mg_ha), "not sampled"),
                 "SD among units" = f1(b$sd_Mg_ha), check.names = FALSE),
      caption = "By stratum. The area mean weights each sampled stratum by its area.")
}
```

```{r all}
ad <- r$all_depths
tab(data.frame(Depth = ad$depth, "Mean (Mg C/ha)" = f1(ad$mean_Mg_ha), "Total (Mg C)" = f0(ad$total_Mg_C),
               "Interval (Mg C/ha)" = ifelse(is.na(ad$ci_low_Mg_ha), "—", paste0(f1(ad$ci_low_Mg_ha), "–", f1(ad$ci_high_Mg_ha))),
               "% estimated" = f0(ad$pct_estimated), check.names = FALSE),
    caption = "All standard depths. The deeper the depth, the more of the number is estimated rather than measured.")
```

**What the interval includes.** `r if (!is.null(res$ci_Mg_ha)) paste("It describes how much the mean could differ because a different set of sampling units might have been drawn.", "It does not include laboratory error, error in the LOI-to-carbon conversion, error in the estimated part below the cores, error in the boundary or stratum areas, or bias from cores that did not recover the full profile.") else "No interval is given for this design. A confidence interval assumes the sampling units were placed at random; these were not, so an interval calculated from them would not tell you how close the mean is to the area's true value."`

```{r fig-map, out.width = "65%"}
fig("area_map.png")
```

The map shows the reporting area, where the cores were taken and one value per area or
stratum. Nothing is mapped between the points: the estimate is an area average, not a map of
where the carbon is.

## 6. Notes from the calculation

```{r notes, results = "asis"}
for (n in res$notes) cat("- ", n, "\n", sep = "")
```

## 7. Missing data, exclusions and limitations

```{r excl, results = "asis"}
bad <- r$checks$cores
if (nrow(bad)) cat("- Left out because they did not pass the checks:", paste(sprintf("%s (%s)", bad$core_id, bad$status), collapse = "; "), "\n") else
  cat("- No cores were left out by the checks.\n")
if (!is.null(res$excluded_area_ha) && res$excluded_area_ha > 0)
  cat(sprintf("- %.2f ha in unsampled strata is excluded: the total covers %.2f ha, not the whole boundary.\n",
              res$excluded_area_ha, res$area_ha))
cat(sprintf("- %s%% of the %g cm mean is estimated below the base of the cores, not measured.\n", f0(res$pct_estimated), D))
cat("- The estimated part assumes carbon below each core behaves like the carbon just above its base (or follows the fitted curve). Deeper cores are the only way to test that.\n")
if (any(r$chk$cores$compaction_basis == "assumed none"))
  cat("- Compaction was not measured on some cores and was assumed to be none.\n")
if (isTRUE(S$BOUNDARY_IS_HYPOTHETICAL)) cat("- The boundary is hypothetical, so the area and total describe no real place.\n")
```

## 8. What this supports, and what would help

```{r supports, results = "asis"}
if (res$design == "exploratory") {
  cat("**Supports:** a first, descriptive figure for the average sediment organic carbon in the area and its",
      "approximate total, stated with its depth and how much of it was estimated.\n\n",
      "**Does not support:** a confidence interval, a claim about precision, a comparison over time, or",
      "any statement about carbon being removed from the atmosphere.\n\n",
      "**What would help:** cores placed with a random or stratified random design (Part 2), enough of them",
      "to meet the precision target, cores long enough to measure the reporting depth, and a mapped boundary.\n")
} else {
  cat("**Supports:** an estimate of the mean and total sediment organic carbon stock in the sampled area to",
      sprintf("%g cm,", D), "with an interval for sampling variation.\n\n",
      "**Does not support:** a map of where carbon is within the area, a rate of accumulation, or removal",
      "of carbon from the atmosphere. The interval leaves out the sources listed in section 5.\n\n",
      "**What would help:** cores that reach the reporting depth (to shrink the estimated share),",
      "elemental organic carbon on a subset of slices to check the LOI conversion, and measured stratum areas.\n")
}
```
