---
title: "Option A — our cores, and how they compare"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  results: "outputs/example/option_A/results.rds"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
r <- readRDS(params$results)
P <- r$project
fig <- function(f) knitr::include_graphics(file.path(r$figures, f))
tab <- function(x, ...) knitr::kable(x, row.names = FALSE, ...)
f1 <- function(x) formatC(x, format = "f", digits = 1)
complete <- r$chk$cores$status == "Complete"
n_cores <- sum(complete)
n_plots <- length(unique(r$chk$cores$plot_id[complete]))
headline <- names(r$refs)[1]
ref1 <- r$refs[[1]]
loi_share <- mean(r$chk$slices$carbon_type[r$chk$slices$check == "OK"] == "LOI")
compaction <- table(r$chk$cores$compaction_basis[complete])
```

**`r r$project$title`**

> **What this report is.** A description of the cores that were measured, and where they sit
> among published eelgrass cores. Each core describes the place it was taken. This report does
> not estimate the carbon in a whole meadow — that needs a defined area and a sampling design
> (Option B).

## 1. The question

`r P$question_A`

## 2. What was sampled

| | |
|---|---|
| Area and habitat | `r P$area_habitat` |
| Survey dates | `r P$survey_dates` |
| How the cores were placed | `r P$design` |
| Cores in this report | `r n_cores` complete core(s), from `r n_plots` plot(s) |
| Data source | `r P$data_source` |
| Prepared by | `r P$prepared_by` |

Each slice is one measurement *within* a core, not an independent sample of the meadow. Cores
sharing a plot are not independent of each other either.

## 3. Carbon pool and depth

**Sediment organic carbon** only — not living eelgrass, roots, or total ecosystem carbon.
Stocks are given for the standard depth increments (0–15, 15–30, 30–50, 50–100 cm) on in-situ
depths, and only where a core reached the bottom of the increment. Cores are compared over
**0–`r r$D` cm**, the deepest standard depth every complete core reached.

## 4. How the numbers were calculated

- **Methods:** `r P$methods`.
- **Slice stock** (Mg C/ha) = dry bulk density (g/cm³) × organic carbon (%) ÷ 100 × slice
  thickness in the tube (cm) × 100. Bulk density = whole-slice dry mass ÷ slice volume, from the
  corer's internal diameter (`r f1(r$chk$corer_diameter_cm)` cm on the Instructions tab
  unless a core gives its own).
- **Organic carbon:** `r if (loi_share > 0) sprintf("%d%% of slices came from loss on ignition, converted with OC%% = %s + %s × LOI%%. The rest are elemental organic carbon.", round(100 * loi_share), r$chk$loi[["intercept"]], r$chk$loi[["slope"]]) else "every slice is elemental organic carbon; no LOI conversion was used."` Total carbon is never used as organic carbon.
- **Compaction:** `r paste(sprintf("%s for %d core(s)", names(compaction), as.integer(compaction)), collapse = "; ")`.
  Compaction moves each slice to its in-situ depth; it never changes how much carbon the slice holds.
- **Increments:** a slice that crosses an increment boundary is split in proportion to its
  overlap, so the increments add up to exactly the carbon measured.
- Nothing is interpolated, smoothed or extrapolated in this report.

## 5. Checks

```{r checks}
st <- r$chk$cores[, c("core_id", "plot_id", "n_slices", "n_ok", "status")]
names(st) <- c("Core", "Plot", "Slices", "Slices OK", "Status")
tab(st)
if (nrow(r$checks$slices)) tab(r$checks$slices, caption = "Slices that did not pass")
if (nrow(r$checks$warnings)) tab(r$checks$warnings, caption = "Worth a look (did not stop a core being totalled)")
```

The R workflow re-calculates every slice independently of the workbook. Largest difference from
the workbook's own numbers:

```{r crosscheck}
cc <- r$cross_check
counts <- grepl("count", cc$quantity)
cc$max_abs_difference <- ifelse(is.na(cc$max_abs_difference), "not available (workbook not recalculated)",
  ifelse(counts, as.character(cc$max_abs_difference),
  ifelse(cc$max_abs_difference < 1e-6, "agrees (< 0.000001)", signif(cc$max_abs_difference, 2))))
names(cc) <- c("Quantity", "Largest difference")
tab(cc)
```

## 6. What the cores show

```{r coretable}
inc <- r$inc
inc$value <- ifelse(inc$status == "complete", f1(inc$stock_Mg_ha),
                    ifelse(inc$status == "partial", "partial", "not reached"))
w <- reshape(inc[, c("core_id", "increment", "value")], idvar = "core_id", timevar = "increment",
             direction = "wide")
names(w) <- sub("^value\\.", "", names(w))
cs <- r$cores[r$cores$status == "Complete", c("core_id", "measured_to_insitu_cm", "bd_g_cm3", "oc_pct")]
cs <- merge(cs, w, by = "core_id")
cs$measured_to_insitu_cm <- f1(cs$measured_to_insitu_cm)
cs$bd_g_cm3 <- formatC(cs$bd_g_cm3, format = "f", digits = 2)
cs$oc_pct <- formatC(cs$oc_pct, format = "f", digits = 2)
names(cs)[1:4] <- c("Core", "Measured to (cm, in situ)", "Bulk density (g/cm³)", "Organic carbon (%)")
tab(cs, caption = "Organic carbon stock by standard increment (Mg C/ha). 'partial' = the core ends inside the increment, so its value is not comparable.")
```

```{r fig-profiles, out.width = "100%"}
fig("profiles.png")
```

```{r fig-increments, out.width = "70%"}
fig("increments.png")
```

```{r fig-bd, out.width = "60%"}
fig("bd_vs_oc.png")
```

A slice far outside the grey cloud is worth re-checking before anything else: a wrong corer
diameter, or a subsample's dry mass set against the whole slice's volume, both show up here first.

## 7. How the cores compare with published eelgrass cores

**What the reference distribution is.** The grey points are individual *Zostera marina* cores
from `r r$region`, compiled by Janousek et al. (2025), measured continuously to at least
`r r$D` cm. They come from different studies, estuaries and years, chosen by their authors
for their own purposes. They show the range other eelgrass meadows have produced — they are not
a random sample of all eelgrass, and they are not a target.

To keep the comparison fair, the reference set leaves out: your own estuary
(`r paste(r$settings$REFERENCE_EXCLUDE_ESTUARIES, collapse = ", ")`); any core within
`r r$settings$REFERENCE_EXCLUDE_WITHIN_M` m of one of your cores (so a published core used as
practice data can never be compared with itself); and any slice the synthesis marks as
estimated rather than measured.

`r if (length(r$refs) > 1) "Two reference sets are shown, because carbon was measured in two ways:" else "One reference set is shown: published cores with elemental organic carbon from studies that removed inorganic carbon."`

```{r refsets}
rt <- r$ref_table
rt[, 5:9] <- lapply(rt[, 5:9], f1)
names(rt) <- c("Reference set", "Cores", "Estuaries", "Studies", "Min", "Lower quartile", "Median",
               "Upper quartile", "Max")
tab(rt, caption = sprintf("Organic carbon stock, 0–%g cm (Mg C/ha)", r$D))
```

The headline set is **`r r$ref_labels[[headline]]`**. `r if (headline == "oc_or_loi") "Its LOI values were converted with the same equation as your own LOI slices, so both sides of the comparison share the same conversion — but the conversion was fitted on your samples, and the published sediments may relate LOI to carbon differently. The measured-only set avoids that, but it is small." else if (length(r$refs) == 1) "No LOI equation was entered on the workbook, so published LOI values are not converted and only directly measured organic carbon is compared." else "It uses only directly measured organic carbon, so no conversion is involved; the larger LOI-converted set is shown for context."`

```{r cmp}
cm <- r$cmp
cm$stock_Mg_ha <- ifelse(is.na(cm$stock_Mg_ha), "", f1(cm$stock_Mg_ha))
pc <- grep("^percentile_", names(cm))
short <- c(oc_or_loi = "OC + converted LOI", oc_measured = "measured OC only")
names(cm) <- c("Core", sprintf("Stock 0–%g cm (Mg C/ha)", r$D), "Note",
               paste0("Percentile — ", short[sub("^percentile_", "", names(cm)[pc])],
                      ifelse(seq_along(pc) == 1, " (headline)", "")))
tab(cm)
```

A percentile is the share of reference cores with a stock at or below yours. It describes where
a core sits; it is not a statistical test, and with a small reference set it moves in large
steps.

```{r fig-cmp, out.width = "100%"}
fig("comparison.png")
```

```{r fig-maps, out.width = "48%", fig.show = "hold"}
fig("locations.png"); fig("reference_map.png")
```

```{r selection}
sel <- attr(ref1, "selection")
tab(data.frame(Step = c("Zostera cores in the synthesis", "In the chosen region", "After excluding your estuary",
                        sprintf("After excluding cores within %g m of yours", r$settings$REFERENCE_EXCLUDE_WITHIN_M),
                        sprintf("Measured continuously to %g cm on this carbon basis", r$D)),
               Cores = as.integer(sel)), caption = "How the headline reference set was chosen")
```

## 8. Missing data, exclusions and limitations

- Cores not in the tables above did not pass the checks in section 5, or did not reach
  `r r$D` cm.
- `r n_assumed <- sum(r$chk$cores$compaction_basis[complete] == "assumed none"); if (n_assumed) sprintf("Compaction was not measured for %d of %d cores and was assumed to be none; if those cores were compacted, their slices sit deeper than shown.", n_assumed, n_cores) else "Compaction was measured on every core."`
- `r if (loi_share > 0) sprintf("%d%% of slices are LOI converted with a fitted equation. That conversion has its own error, which is not shown in any number here.", round(100 * loi_share)) else "All slices are elemental organic carbon."`
- Reference-core depths are as published, not corrected for compaction.
- Stocks below the base of a core are unknown here. A core's total is only comparable with
  another core's over the same depth.

## 9. What this supports, and what would help

**Supports:** a description of sediment organic carbon at the places these cores were taken,
over 0–`r r$D` cm, and where those values sit among published eelgrass cores from the region.

**Does not support:** a meadow-wide average, a total for an area, an interval around a mean, or
any statement about carbon being added (sequestration) — that needs dated cores.

**What would help:** more cores, placed with a sampling design (Part 2), to describe the meadow
rather than single points; deeper cores to compare over 0–30 cm or more; elemental organic
carbon on a subset of slices to check the LOI conversion; measured compaction on every core.
