---
title: "Eelgrass Carbon Stock"
date: today
engine: knitr
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
# This report RENDERS the pipeline — it does not reimplement it. Every number
# below comes from the same scripts you run with source("run_pipeline.R"), so
# the report and the console can never disagree.
library(dplyr); library(ggplot2); library(knitr)
invisible(capture.output(suppressMessages({
  source("run_pipeline.R")
})))
mg <- function(x) x * 10          # kg C/m² -> Mg C/ha
site_total <- function(e) e$estimate * sum(STRATUM_AREAS_M2) / 1000
```

**Site:** `r SITE_NAME`

## The headline

```{r headline}
data.frame(
  Quantity = c(
    sprintf("Carbon stock to %d cm (measured)", PRIMARY_DEPTH_CM),
    sprintf("Carbon stock to %.0f cm (partly modelled)", DEEP_DEPTH_CM),
    "After combining with the regional prior",
    sprintf("Site total to %d cm", PRIMARY_DEPTH_CM),
    "Precision achieved",
    "Precision target"),
  Value = c(
    sprintf("%.1f ± %.1f Mg C/ha", mg(est_primary$estimate), mg(est_primary$se)),
    sprintf("%.1f ± %.1f Mg C/ha", mg(est_deep$estimate),    mg(est_deep$se)),
    sprintf("%.1f ± %.1f Mg C/ha", posterior$mean, posterior$sd),
    sprintf("%.0f Mg C over %.1f ha", site_total(est_primary),
            sum(STRATUM_AREAS_M2) / 10000),
    sprintf("±%.1f%% at %.0f%% confidence", 100 * (est_primary$ci[1,2] -
            est_primary$estimate) / est_primary$estimate, 100 * CONF_LEVEL),
    sprintf("±%.0f%% — %s", 100 * TARGET_MARGIN,
            if (posterior$passes) "met" else "NOT met"))
) |> kable(align = "lr")
```

`r if (!posterior$passes) sprintf("The campaign did not reach its precision target. That is a result, not a failure — it says the interval is too wide to manage by, and the analysis below shows what would close it.")`

## What was collected

`r nrow(core_totals)` cores across `r n_distinct(core_totals$stratum)` strata, `r nrow(cores)` sediment slices.

```{r cores-table}
core_totals |>
  transmute(Core = core_id, Stratum = stratum,
            `Stock 0-25 cm (kg C/m²)` = round(stock_primary, 2),
            `Stock 0-50 cm (kg C/m²)` = round(stock_deep, 2),
            `% modelled` = round(pct_modelled, 1)) |>
  kable(align = "llrrr")
```

```{r profiles, fig.height=4.5}
p_soc
```

## Compaction

A percussion corer compresses sediment: you recover a shorter core than the
hole you made. Correcting for it restores the original depths — and must leave
the carbon unchanged, because the tube holds exactly what came out of the
ground. The depths stretch by the compaction factor and the bulk density
thins by the same factor, so the two cancel.

```{r compaction}
cores |>
  distinct(core_id, compaction_factor) |>
  left_join(cores |> group_by(core_id) |>
              summarise(base = max(depth_bottom_cm), .groups = "drop"),
            by = "core_id") |>
  transmute(Core = core_id,
            `Compaction factor` = round(compaction_factor, 3),
            `In-situ depth (cm)` = round(base, 1)) |>
  kable(align = "lrr")
```

## Putting cores on common depths

Cores reach different depths and were sliced at different thicknesses, so they
are splined onto standard intervals with a mass-preserving spline
(`mpspline2`). Below the base of a core the profile is **modelled**, not
measured, using an asymptotic curve with a non-zero recalcitrant floor.

Every interval records how much of it is measured rather than modelled, and
stocks are apportioned by that fraction — so a core ending at 26 cm contributes
only its real 4% to the 25–50 cm interval.

```{r harmonized, fig.height=5}
p_harmonized
```

The deep figure is reported to **`r sprintf("%.0f", DEEP_DEPTH_CM)` cm**, the
deepest depth every core supports. Cores that went deeper are not credited for
it; that is the price of one comparable number.

## The estimate

Each stratum's mean is weighted by its **area**, not by how many cores landed
in it, using a design-based stratified estimator (`survey`) with a finite
population correction.

```{r estimate, fig.height=4}
p_stratified
```

```{r strata-table}
as.data.frame(by_stratum) |>
  transmute(Stratum = STRATUM_LABELS[stratum],
            `Mean 0-25 cm (kg C/m²)` = round(stock_primary, 2),
            `SE` = round(se.stock_primary, 2)) |>
  kable(align = "lrr", row.names = FALSE)
```

The interval rests on **`r degf(design)` degrees of freedom**. With this few
cores it is wide, and no amount of analysis narrows it — only more cores, or a
looser target, will.

## Combining with what was already known

The planning stage used a prior to size the campaign. That prior is still true
afterwards, so it is used again here as evidence — a weighted average, where
each source contributes in proportion to its precision.

The prior comes from `r PRIOR_N` *Zostera marina* cores across
`r PRIOR_N_ESTUARIES` estuaries in the Janousek et al. (2025) synthesis,
filtered to sites that resemble this one and **excluding this site's own
cores**, which appear in that synthesis and would otherwise be counted twice.

```{r update, fig.height=4.5}
p_update
```

```{r update-table}
data.frame(
  Source = c("Prior (sites like this one)", "Evidence (these cores)",
             "Posterior (combined)"),
  `Mg C/ha` = c(sprintf("%.1f ± %.1f", PRIOR_MEAN, PRIOR_SD_MEAN),
                sprintf("%.1f ± %.1f", mg(est_primary$estimate), mg(est_primary$se)),
                sprintf("%.1f ± %.1f", posterior$mean, posterior$sd)),
  Weight = c(sprintf("%.0f%%", 100 * posterior$prior_weight),
             sprintf("%.0f%%", 100 * (1 - posterior$prior_weight)), "—"),
  check.names = FALSE) |> kable(align = "lrr")
```

The prior earned **`r sprintf("%.0f%%", 100 * posterior$prior_weight)`** of the
weight. `r if (posterior$prior_weight < 0.2) "Most of the answer comes from the fieldwork, which is what a campaign of this size should produce."`

## What to be careful about

- **The deep number is partly modelled.** Below each core's base the profile is
  an extrapolation, and the per-core modelled share is in the table above.
- **`r degf(design)` degrees of freedom** is very few. The confidence interval
  is honest but wide, and the between-stratum comparison has low power — a
  non-significant result there is not evidence the strata are equal.
- **Stratum areas drive the site total.** They come from `STRATUM_AREAS_M2` in
  `00_config.R`; if those are placeholders, so is the total.
- **The prior assumes independence.** It excludes this site's own cores. If you
  move this workflow to another meadow, check `PRIOR_EXCLUDE_ESTUARY` again.
- **No map here.** Mapping where the carbon sits needs far more cores than
  this — see `06_advanced_spatial.R`.

## Methods

```{r methods}
data.frame(
  Step = c("Compaction", "Depth harmonization", "Below-core profile",
           "Stratified estimation", "Prior", "Combination"),
  Method = c("Depth stretch with matching bulk-density correction",
             "Mass-preserving (equal-area) spline",
             "Asymptotic model with a non-zero recalcitrant floor",
             "Design-based stratified mean with finite population correction",
             "Similarity-matched subset of a regional core synthesis",
             "Normal-Normal precision-weighted update"),
  Source = c("from scratch — matches the digital data sheet",
             "**mpspline2** — Bishop et al. (1999); Malone et al. (2009)",
             "base **stats** `SSasymp` — shape after Kindeberg et al. (2019)",
             "**survey** — Lumley (2004, 2010); Cochran (1977)",
             "Janousek et al. (2025), figshare 28127486",
             "from scratch — Gelman et al. (2013) sec. 2.5")
) |> kable(align = "lll")
```

---

*Generated `r format(Sys.Date())` from the pipeline in this folder. Every
number above is produced by `run_pipeline.R`; this report contains no separate
calculation.*
