---
title: "Going further, module 4 — what published cores add"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  results: "../../outputs/example/module4/results.rds"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
r <- readRDS(params$results)
fig <- function(f) knitr::include_graphics(file.path(r$figures, f))
tab <- function(x, ...) knitr::kable(x, row.names = FALSE, ...)
f1 <- function(x) ifelse(is.na(x), "—", formatC(x, format = "f", digits = 1))
pc <- function(x) ifelse(is.na(x), "—", paste0(round(100 * x), "%"))
t <- r$tab; pr <- r$prior; lo <- r$loeo_tab
comb <- t[t$source == "Combined", ]
```

**`r r$project$title`**

> **An optional extra, not part of Options A or B.** This is a *model-based* estimate: it
> assumes your sampling units behave like a random draw from your meadow, and your meadow like
> one more estuary from the published set. It does not replace the Option B estimate, and it is
> no better than those assumptions.

## 1. The question

With only a few cores, how much do published eelgrass cores from other estuaries sharpen our
estimate of the meadow's average sediment organic carbon over **0–`r r$D` cm**?

## 2. The two sources

**Published cores.** `r pr$n_cores` *Zostera marina* cores from `r pr$n_estuaries` estuaries
(Janousek et al. 2025), measured to at least `r r$D` cm, carbon basis: *`r r$ref_label`*. Your
own estuary (`r paste(r$settings$REFERENCE_EXCLUDE_ESTUARIES, collapse = ", ")`) and any core
within `r r$settings$REFERENCE_EXCLUDE_WITHIN_M` m of yours are left out, so nothing is
counted twice. Estuaries differ more than cores within an estuary do: estuary-to-estuary
differences make up `r round(100 * pr$icc)`% of the variation. This spread is what keeps the
prior wide.

**Your cores.** `r nrow(r$units)` sampling unit(s) (`r r$settings$SAMPLING_UNIT`s), measured
to `r r$D` cm. Only measured carbon is used here; nothing is extrapolated.

```{r fig-prior, out.width = "75%"}
fig("prior_estuaries.png")
```

## 3. The result

```{r result}
x <- data.frame(Source = t$source, "Mean (Mg C/ha)" = f1(t$estimate_Mg_ha),
                "Interval (Mg C/ha)" = ifelse(is.na(t$low_Mg_ha), "—", paste0(f1(t$low_Mg_ha), "–", f1(t$high_Mg_ha))),
                "Weight given to the published cores" = pc(t$weight_of_prior), check.names = FALSE)
tab(x, caption = sprintf("%g%% intervals, 0–%g cm", 100 * r$conf, r$D))
```

```{r fig-borrow, out.width = "90%"}
fig("borrow_strength.png")
```

The combined estimate is a weighted compromise between the two sources. The published cores
received **`r pc(comb$weight_of_prior)`** of the weight; your cores received the rest. The
interval for "your cores only" uses the spread among your own few cores, which is itself
uncertain with so few. The combined interval instead uses how much cores typically differ within
a meadow in the published set.

## 4. Does it work? Checked on the published estuaries themselves

Each reference estuary in turn played "your meadow". The prior was rebuilt without it, a few of
its cores were drawn at random (200 times), and both methods were compared with the mean of the
cores *not* drawn.

```{r loeo}
tab(data.frame("Cores taken" = lo$k, "Estuaries tested" = lo$estuaries,
               "Combined: average error (Mg C/ha)" = f1(lo$combined_mean_abs_error),
               "Combined: interval covered the truth" = pc(lo$combined_coverage),
               "Local only: average error (Mg C/ha)" = f1(lo$local_mean_abs_error),
               "Local only: interval covered the truth" = pc(lo$local_coverage), check.names = FALSE),
    caption = sprintf("Target coverage is %g%%. With 0 cores the 'combined' row is the prior alone; with 1 core no local interval can be made.", 100 * r$conf))
```

```{r fig-loeo, out.width = "65%"}
fig("leave_one_estuary_out.png")
```

`r k <- lo$k[!is.na(lo$local_mean_abs_error)]; better <- k[lo$combined_mean_abs_error[!is.na(lo$local_mean_abs_error)] < lo$local_mean_abs_error[!is.na(lo$local_mean_abs_error)]]; if (length(better) == length(k)) "Here, combining was closer than local cores alone at every number of cores tested." else if (length(better)) sprintf("Here, combining was closer than local cores alone with %s cores, but not with %s.", paste(better, collapse = " and "), paste(setdiff(k, better), collapse = " and ")) else "Here, combining was not closer than local cores alone at any number of cores tested."`

**Reading this check honestly.** It uses only `r max(lo$estuaries)` estuaries, and fewer at the
larger core counts, so it is a rough guide. "Truth" here is the mean of a few unseen cores. That
is itself noisy, so coverage is expected to fall a little below the target even for a sound
method. Coverage well below `r 100 * r$conf`% means the intervals are too narrow to rely on.

## 5. What this does not include

- Error in the LOI-to-carbon conversion, laboratory error, and compaction uncertainty.
- Uncertainty in the prior's own spread, which is estimated from a handful of estuaries.
- Any test of whether your meadow really resembles the published ones. Differences in sediment,
  exposure or depth can make it unlike all of them.

## 6. What it supports

**Supports:** a second, model-based view of your meadow's average stock over 0–`r r$D` cm,
showing how much a small set of cores can be informed by published data.

**Does not support:** an area total on its own (use Option B for the area and the design), any
claim of crediting-grade precision, or any statement about carbon accumulation. Modules 5 and 6
(see the Going further README) take the same idea further. Each adds assumptions that need
their own checks.
